---
title: "Dealing with different identifications of the same animal"
output: html_notebook
---

Because we host our project on the zooniverse, it is possible to have multiple people score each event, which means we may not get the same answers for each event in terms of species id, behavior, etc.

To that end, we need a way to check which events have differences in their identifications of species and of e.g. behaviors.  Once those events are identified, we need an efficient way to go back to those events and make a final appraisal of their contents.

The following bunch of R scripts takes the flattened classfication file produced by running the *Flatten-zooniverse-NoCoWild.Rmd* scripts and does the following:  

1. Identify which events have more than one classification
2. For those events with more than 1 classification, determine which ones have more than one species identified.

```{r}
#set up workspace
rm(list = ls())
library(dplyr)
library(ggplot2)
```
```{r}
#pull in the data file
classification_file<-"sampledatasmall.csv-flattened.csv"
library(readr)
DF<- read_csv(classification_file)
```

Now let's find the repeats
```{r}
Repeats<-DF %>% group_by(subject_ids) %>% filter(n()>1)
Repeats

#gives those subject ids that appear more than once
```
Now see if I can get a list of species ids for each subject that was classified more than once.  Try looping through the data by subject id's and putting answers in a new DF

First, determine how many there are of each subject id

```{r}
numSubjects <- DF %>% group_by(subject_ids) %>% summarise(
	numObs = length(subject_ids)
)

#and get the ids of those with >1 classification

repeatSubs<-filter(numSubjects, numObs > 1)

#why do I get a different number? Because the summary doesn't include both cases of the repeat, but the Repeats list does.

#so repeatSubs is the list of subjects that repeat > 1 time.  Can use it to loop if needed.
```

